Graph Learning from Binomial Signals

### 2022-05-26 Documents/paper/draft.Rmd
Draft update.

---
title: 'Draft: Learning Laplacian from Binomial Signals'
author: "Skip Moses"
date: "3/12/2022"
output:
  bookdown::pdf_document2: default
  html_document:
    df_print: paged
cls: IEEEtran.cls
bibliography: ref.bib
---

@@

@@

Dong et al. propose a modified Factor Analysis model in @dong2016learning for learning a valid graph laplacian. The model proposed is given by 
$$x = \chi h + \mu_x + \epsilon$$
where $h \in \mathbb{R}^N$ is the latent variable, $\mu_x \in \mathbb{R}^N$ mean of $x$. The noise term $\epsilon$ is assumed to follow a multivariate Gaussian distribution with mean zero and covariance $\sigma_\epsilon^2I_N$. The key difference from traditional factor analysis is the choice of $\chi$ as a valid eigenvector matrix for a graph laplacian. Finding a maximum apriori estimate of $h$ reduces to solving the optimization problem

@@

Both (\ref{eq:dongprob1}) and (\ref{eq:dongprob2}) can be cast as convex optimization problems. Specifically, \ref{eq:dongprob1} can be solved with the method of alternating direction of multipliers (ADMM). The model is applied to both synthetic and real world data, and compared to a technique used in machine learning that is similar to sparse inverse covariance estimation of Gaussian Markov Random Field models. Four evaluation criteria were used  to evaluate the performance of the framework: $\textit{Precision, Recall, F-measure,}$ and $\textit{Normalized Mutual Information (NMI)}$. Kalofolias cites two weaknesses to this model: The use of the Frobenius norm is not amenable to interpretation, and it requires four constraints on the solution space @kalofolias2016learn.

@@

The model proposed by Kalofolias is given by specifying $f(A) = -\alpha \textbf{1}^T\log(W\textbf{1}) + \beta||W||_F^2$. An appealing feature of this model is the fact that for any $\lambda >0$ and solution $F(Z,\alpha, \beta)$ the following equalities hold
\begin{equation}
	\label{eq:KalofoliasEquality}
	F(Z, \alpha, \beta) = \lambda F(Z, \frac{\alpha}{\lambda}, \beta \lambda) = \alpha F(Z, 1, \alpha \beta).
\end{equation}
This means we can fix a scale $||A|| = s$ for an arbitrary norm and solve for $\beta$ in \ref{eq:kalotransform} with $\alpha = 1$, and then re-normalize our solution. Furthermore, the log barrier term only acts on the degrees. This provides an advantage over the model \ref{eq:dongmodel}, because it promotes connectivity in the learned graph. Kalofolias goes onto show \ref{eq:kalotransform} can be solved efficiently for many choices of $f(A)$ via primal dual algorithms.

In @venkitaraman2019predicting machine learning concepts are employed to learn a graph from training data. The goal is given some data ${x_n}$ to model a target $t_n$ by the linear basis model for regression
\begin{equation}
	\label{eq:LinearModel}
	y_n = \textbf{W}^T\phi(x_n),
\end{equation}
where $\phi(x_n)$ is some known function and $\textbf{W}$ denotes the regression coefficient matrix. Venkitarman et al. assume only the target $y_n$ is smooth over the graph. The input $x_n$ does not have to be a signal over the graph and can even be agnostic to a graph. The cost function for the model when the underlying graph is known is given by 
\begin{equation}
	\label{eq:cost}
	C(\textbf{W}) = \sum_{n=1}^N\left( ||t_n - y_n||_2^2 + \beta y_n^TLy_n \right) + \alpha \text{tr}(\textbf{W}^T\textbf{W}).
\end{equation}
Next, it is shown KRG induces a smoothing effect on the target by a shrinkage in the direction of eigenvectors of $L$ that have eigenvalues less than $\alpha$. In order to learn an unknown graph, a term is added to the cost function $C(\textbf{W}, L) = C(\textbf{W}) + \nu \text{tr}(L^TL)$.  Venkitarman et al. use the model to predict temperature using air-pressure observations collected in Sweden, next day temperature from current day in Sweden, and fMRI voxel intensities of the cerebellum region. The latter two experiments are compared the method of kernel-ridge regression (KRR) by using a normalized-mean-square-error to measure the performance of the prediction. It is concluded that the method is well suited for training with noise data sets that may be small.

@@

Therefore, our Likelihood function will be given by

@@

\begin{equation}
\label{eq:optimization}
\begin{aligned}
&\max_{\mathbf{A},h} \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(\mathbf{A}[i,]h) - \log(1 + e^{\mathbf{A}[i,]h})\right) - \alpha \vert L \vert _ F \\
\textrm{s.t.} \quad  &\mathbf{A}_{i,j} = 0 \text{ if } i =j;\,\, \mathbf{A}_{i,j} \geq 0 \text{ if } i \neq j\\
                     &\max(h) \leq a ;  \,\, \min(h) \geq b \,\, ; \mathbb{1}h^T = 0  
\end{aligned}
\end{equation}
where $\alpha$ is a tuning parameter for controlling the sparsity and $\mathbf{A}$ and $b$ are tuning parameters for restricting the spread of the values of $h$.

@@

<hr />
\begin{align*}
&\textbf{Algorithm 1: Binary Signal Graph Learning} \\
&1) \textbf{ Input: } \text{ Input a signal } \mathbf{Y}.\\
&2) \textbf{ Output: } \text{ Output an estimated } \mathbf{A}.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } \mathbf{A} \textbf{ given } h:\\
&6) \,\,\,\, \textit{ Fix }  h \textit{ in Optimization Program and solve } \mathbf{A} \\
&7) \textbf{        Update } h \textbf{ given } \mathbf{A}:\\
&8) \,\,\,\, \textit{ Fix } \mathbf{A} \text{ in Optimization Program and solve h} \\
&9) \textbf{ end for }
\end{align*}
<hr />

@@

Let $\mathbf{A}$ be the adjacency matrix of the graph in Figure \@ref(fig:GTG), and $h \in \mathbb{R}$ be such that $h_i \sim \mathcal{N}(0,1)$. Set $p$ as a logistic function of $\mathbf{A}h$, and compare $p$ to a vector $t$, where each $t_i$ follows a random uniform distribution $\mathcal{U}(0,1)$. If $t_k < p_k$, then node $k$ takes the signal value of $1$, and $0$ otherwise. We generate $M = 100$ synthetic signals, and use them to learn an adjacency matrix.

@@

The learned graph is found with parameter values $\alpha = .2$, $a = 1$ and $b = -1$ by implementing Algorithm 1. We can see the learned adjacency matrix shares some similar features to the ground truth (left), but the learned adjacency matrix is sparser than the ground truth (right).

@@

\newpage
# References

>>>>

---
title: 'Graph Learning from Binomial Signals'
author: "Skip Moses"
date: \today
output:
  bookdown::pdf_document2: default
  html_document:
    df_print: paged
cls: IEEEtran.cls
bibliography: ref.bib
---

@@

In this work, we suppose the odds of a node having a signal value of $1$ or $0$ depends on the signal value of its neighboring nodes in order to estimate the underlying graph topology. Our methodology generalizes the framework of traditional logistic regression in order to learn an adjacency matrix from a set of binary signals.

@@

Dong et al. propose a modified factor analysis model in @dong2016learning for learning a valid graph laplacian. The model proposed is given by 
$$x = \chi h + \mu_x + \epsilon$$
where $h \sim \mathcal{N}(0, \Lambda^{\dagger})$ is the latent variable,$\Lambda$ is the eigenvalue matrix of $L$, and  $\mu_x \in \mathbb{R}^N$ mean of $x$. The noise term $\epsilon$ is assumed to follow a multivariate Gaussian distribution with mean zero and covariance $\sigma_\epsilon^2I_N$. The key difference from traditional factor analysis is the choice of $\chi$ as a valid eigenvector matrix for a graph laplacian. Finding a maximum apriori estimate of $h$ reduces to solving the optimization problem

@@

Both (\ref{eq:dongprob1}) and (\ref{eq:dongprob2}) can be cast as convex optimization problems. Specifically, \ref{eq:dongprob1} can be solved with alternating direction method of multipliers (ADMM). The model is applied to both synthetic and real world data, and compared to a technique used in machine learning that is similar to sparse inverse covariance estimation of Gaussian Markov Random Field models. Four evaluation criteria were used  to evaluate the performance of the framework: $\textit{Precision, Recall, F-measure,}$ and $\textit{Normalized Mutual Information (NMI)}$. Kalofolias cites two weaknesses to this model: The use of the Frobenius norm on the Laplacian  has reduced interpretability, and it requires four constraints on the solution space @kalofolias2016learn. These extra constraints make computational scalability difficult.

@@

The model proposed by Kalofolias is given by specifying $f(A) = -\alpha \textbf{1}^T\log(A\textbf{1}) + \beta||A||_F^2$. An appealing feature of this model is the fact that for any $\lambda >0$ and solution $F(Z,\alpha, \beta)$ the following equalities hold
\begin{equation}
	\label{eq:KalofoliasEquality}
	F(Z, \alpha, \beta) = \lambda F(Z, \frac{\alpha}{\lambda}, \beta \lambda) = \alpha F(Z, 1, \alpha \beta).
\end{equation}
This means we can fix a scale $||A|| = s$ for an arbitrary norm and solve for $\beta$ in \ref{eq:kalotransform} with $\alpha = 1$, and then re-normalize our solution. Furthermore, the log barrier term only acts on the degrees. This provides an advantage over the model \ref{eq:dongmodel}, because it promotes connectivity in the learned graph. Kalofolias goes onto show \ref{eq:kalotransform} can be solved efficiently for many choices of $f(A)$ via primal dual algorithms.

In @venkitaraman2019predicting machine learning concepts are employed to learn a graph from training data. Let $\{x_n\}_1^N$ be a set of $N$ input observations with each $x_n$ paired with a target $t_n \in \mathbb{R}^M$. The authors model a set of smooth signals $y_n$ and $t_n$ via the following: 
\begin{equation}
	\label{eq:LinearModel}
	y_n = \textbf{W}^T\phi(x_n),
\end{equation}
where $\phi(x_n)$ is some known function and $\textbf{W}$ denotes the regression coefficient matrix. Venkitarman et al. assume only the target $y_n$ is smooth over the graph. The input $x_n$ does not have to be a signal over the graph and can even be agnostic to a graph. The cost function for the model when the underlying graph is known is given by 
\begin{equation}
	\label{eq:cost}
	C(\textbf{W}) = \sum_{n=1}^N\left( ||t_n - y_n||_2^2 + \beta y_n^TLy_n \right) + \alpha \text{tr}(\textbf{W}^T\textbf{W}).
\end{equation}
Next, it is shown KRG induces a smoothing effect on the target by a shrinkage in the direction of eigenvectors of $L$ that have eigenvalues less than $\alpha$. In order to learn an unknown graph, a term is added to the cost function $C(\textbf{W}, L) = C(\textbf{W}) + \nu \text{tr}(L^TL)$.  Venkitarman et al. use the model to predict temperature using air-pressure observations collected in Sweden, next day temperature from current day in Sweden, and fMRI voxel intensities of the cerebellum region. The latter two experiments are compared the method of kernel-ridge regression (KRR) by using a normalized-mean-square-error to measure the performance of the prediction. It is concluded that the method is well suited for training with noise data sets that may be small @venkitaraman2019predicting .

@@

Therefore, our likelihood function will be given by

@@

\begin{equation}
\label{eq:optimization}
\begin{aligned}
\max_{\mathbf{A},h} \phantom{..}& \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(\mathbf{A}[i,]h) - \log(1 + e^{\mathbf{A}[i,]h})\right) - \alpha \vert A \vert _ F \\
\textrm{s.t.}  & \quad \mathbf{A}_{i,j} = 0 \text{ if } i =j \\
               & \quad \mathbf{A}_{i,j} \geq 0 \text{ if } i \neq j\\
               & \quad h \sim \mathcal{U}(-1,1) 
\end{aligned}
\end{equation}
where $\alpha$ is a tuning parameter for controlling the sparsity and $a$ and $b$ are tuning parameters for restricting the spread of the values of $h$.

@@

<hr />
\begin{align*}
&\textbf{Algorithm 1: Binary Signal Graph Learning} \\
&1) \textbf{ Input: } \text{ Input a signal on a graph} \mathbf{Y}.\\
&2) \textbf{ Output: } \text{ Output an estimated } \mathbf{A}.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } \mathbf{A} \textbf{ given } h:\\
&6) \,\,\,\, \textit{ Fix }  h \textit{ in Optimization Program and solve } \mathbf{A} \\
&7) \textbf{        Update } h \textbf{ given } \mathbf{A}:\\
&8) \,\,\,\, \textit{ Fix } \mathbf{A} \text{ in Optimization Program and solve h} \\
&9) \textbf{ end for }
\end{align*}
<hr />

@@

Let $\mathbf{A}$ be the adjacency matrix of the graph in Figure \@ref(fig:GTG), and $h \in \mathbb{R}$ be such that $h_i \sim \mathcal{U}(-1,1)$. Set $p$ as a logistic function of $\mathbf{A}h$, and compare $p$ to a vector $t$, where each $t_i$ follows a random uniform distribution $\mathcal{U}(0,1)$. If $t_k < p_k$, then node $k$ takes the signal value of $1$, and $0$ otherwise. We generate $M = 100$ synthetic signals, and use them to learn an adjacency matrix.

@@

The optimal $\alpha$ value, $\alpha = 0.2$ is found by performing a grid search, and selecting the $\alpha$ value that maximizes the F-measure of the learned adjacency matrix with the ground truth adjacency matrix. The parameter values for $a = 1$ and $b = -1$ where selected to be the ground truth values to simplify parameter optimization. We can see the learned adjacency matrix shares some similar features to the ground truth (left), but the learned adjacency matrix is sparser than the ground truth (right).

@@

\newpage
